An Empirical Analysis of the Internet Engineering Task Force with Computational Methods

### 2024-09-29 paper.tex
Update §4.2 and §4.3

Figure~\ref{fig:author_affiliations_normalised} shows the top 10
affiliations by proportion of authors each year. Affiliation data is
gathered from the Datatracker, and processed to normalise affiliation
names, removing common variations in spelling, and amalgamating known
subsidiaries and merged companies. For example, Huawei and Futurewei are
combined as Huawei, and Sun Microsystems is merged with Oracle.

From Figure~\ref{fig:author_affiliations_normalised}, we observe several
interesting trends.  First, Cisco remains a consistent employer of IETF
contributors, with around 12\% of authors affiliated with the company in
2020, and having been the single largest affiliation across all years in
the dataset. We can also see the rise of Huawei beginning in 2005, with
7.1\% of authors being affiliated with the company in 2020, having peaked
at 9.7\% in 2018. Google has a similar trajectory, first appearing in the
dataset in 2006, with 3.8\% of authors being affiliated with it in 2020.

@@

That we are able to observe changing trends in author affiliations, and in
particular, that participants from new companies contribute, indicates that
the IETF has broader relevance beyond a narrow set of companies. While we
do not demonstrate any causal link, it is encouraging that commercially
successful companies opt to enable their employees to actively participate
in the IETF. Care is needed to ensure that this relevance is maintained,
however: we observe that the author pool has grown less diverse in terms of
companies that are represented. 35.4\% of authors came from the top 10
affiliations in the dataset in 2020, compared with 25.6\% in 2001.

@@

%..................................................................................................
\pb{Academia and consultants:}
Academic affiliations are those where the affiliation name contains
``University'', ``Institute'', or ``College'', while consultancy
affiliations are those that contain ``Consultant``. Affiliation data has
been normalised to remove common abbreviations (e.g., ``U.'' for
``University'') and to translate non-English affiliations.  As shown in
Figure~\ref{fig:author_affiliations_normalised}, we find that an increasing
number of authors come from academic affiliations, growing from 8.1\% of
authors in 2001, to 13.6\% in 2020, having peaked at 16.5\% in 2009. The
number of consultants, by our measure, has remained stable, accounting for
2\% of authors in 2020.  The remaining authors are largely from industrial
affiliations.

Figure~\ref{fig:author_affiliations_normalised_acad} shows the top 10
academic affiliations in the dataset, and the percentage of academic
authors that have those affiliations over time. In general, academic
affiliations are each typically held by a small number of authors. We can
see a number of trends in academic authorship, with fewer authors from
Columbia University, MIT, and ISI in recent years, and the rise of Tsinghua
University and University Carlos III of Madrid.

@@

%..................................................................................................
% IMC 2021 paper Section 3.2 (last part)
\pb{Arrival of new authors:}
Finally, Figure~\ref{fig:author_new} shows the percentage of authors in each
year that have not previously authored an RFC. Given that the dataset used here
begins in 2001, 100\% of authors are new in that year. The more stable trend in
recent years likely highlights the churn in RFC authorship, with around 30\%
of authors each year having never previously authored an RFC.
\todo{Not clear this fits here}

@@

%--------------------------------------------------------------------------------------------------
\subsection{Summary}
The trends highlighted in this section  indicate a pool of authors that is
diversifying and changing over time, with a growing proportion of authors
from outside of North America, contributions from new companies and
affiliations, and relatively high authorship churn. Despite this, we find
that certain regions and groups are not well represented in the IETF (e.g.,
contributors from Africa), and that the authorship pool is becoming
increasingly centralised, with a third of authors coming from the top 10
affiliations. This suggests, if the organisation is concerned with being
more representative, further efforts are needed.

>>>>

To explore trends in author affiliations, we gather affiliation data from
the IETF Datatracker.  This data is processed to normalise affiliation
names by removing common variations in spelling and to amalgamate known
subsidiaries and merged companies. For example, Huawei and Futurewei are
combined as Huawei, and Sun Microsystems is merged with Oracle.

%..................................................................................................
\pb{Corporate Authors:}
Figure~\ref{fig:author_affiliations_normalised} shows the top ten
affiliations by proportion of RFC authors each year. We observe several
interesting trends.  First, Cisco remains a consistent employer of IETF
contributors, with around 12\% of authors affiliated with the company in
2020, and having been the single largest affiliation across all years in
the dataset. We can also see the rise of Huawei beginning in 2005, with
7.1\% of authors being affiliated with the company in 2020, having peaked
at 9.7\% in 2018. Google has a similar trajectory, first appearing in the
dataset in 2006, with 3.8\% of authors being affiliated with it in 2020.

@@

These shifts in affiliation reflect changes in fortune and commercial
success of the companies. They also show the ongoing importance of IETF
standards, with new companies choosing to send participants: while we do
not demonstrate any causal link, it is encouraging that commercially
successful companies opt to enable their employees to actively participate
in the IETF. Care is needed to ensure that this relevance is maintained,
however: we observe that the author pool has grown less diverse in terms of
companies that are represented. 35.4\% of authors came from the top 10
affiliations in the dataset in 2020, compared with 25.6\% in 2001.

@@

%..................................................................................................
\pb{Academia and consultants:}
Academic affiliations are those where the affiliation name contains
``University'', ``Institute'', or ``College'', while consultancy
affiliations are those that contain ``Consultant`` (we recognise that this
definition of consultants is limited and incomplete). Affiliation data has
been normalised to remove common abbreviations (e.g., ``U.'' for
``University'') and to translate non-English affiliations.  As shown in
Figure~\ref{fig:author_affiliations_normalised}, we find that an increasing
number of authors come from academic affiliations, growing from 8.1\% of
authors in 2001, to 13.6\% in 2020, having peaked at 16.5\% in 2009. The
number of consultants, by our limited measure, has remained stable,
accounting for 2\% of authors in 2020.  The remaining authors are largely
from industrial affiliations.

Figure~\ref{fig:author_affiliations_normalised_acad} shows the top 10
academic affiliations in the dataset, and the percentage of academic
authors that have those affiliations over time. In general, academic
affiliations are each typically held by a small number of authors. We can
see a number of trends in academic authorship, with fewer authors from
Columbia University, MIT, and ISI in recent years, and the rise of Tsinghua
University.

While some academic affiliations are shared by large groups of participants
(e.g., MIT, ISI, Tsinghua University), academic participation is often
driven by a single research group or individual within an institution.
This allows smaller institutions, such as the University of Bremen and
the University of Auckland, that appear in Figure~\ref{fig:author_affiliations_normalised_acad},
and others with single but prolific authors, to have outsize influence.
Understanding whether this indicates academics pushing for standardisation
of the results of government funded research to increase its impact, the
outcomes of industry sponsored research, or academics acting as consultants
for industry, is for further study.

@@

%..................................................................................................
% IMC 2021 paper Section 3.2 (last part)
\pb{Arrival of new authors:}
Figure~\ref{fig:author_new} shows the percentage of authors in each year
that have not previously authored an RFC. Given that the dataset used here
begins in 2001, 100\% of authors are new in that year. The more stable
trend in recent years likely highlights the churn in RFC authorship, with
around 30\% of authors each year having never previously authored an RFC.
This is consistent with the presence of authors with new affiliations
and from new countries and shows an ongoing influx of new participants.

@@

%--------------------------------------------------------------------------------------------------
\subsection{Summary}
The trends highlighted in this section  indicate a pool of authors that is
slowly diversifying and changing over time, with a growing proportion of authors
from outside of North America, contributions from new companies and
affiliations, and relatively high authorship churn. Despite this, we find
that certain regions and groups are not well represented in the IETF (e.g.,
contributors from Africa), and that the authorship pool is becoming
increasingly centralised, with a third of authors coming from the top 10
affiliations. This suggests that, if the IETF is concerned with being
more representative, further efforts are needed.
